Analysis of H-1B Visa Workers Using H-1B LCA Disclosure Data (2020–2024)

### 2025-02-21 products/manuscript/Manuscript.qmd
part 2 commit 1

The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

@@

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

@@

# Summary/Abstract
_Write a summary of your project._

@@

The data is H1B LCA Disclosure Data (2020-2024) from [Kaggle](https://www.kaggle.com/datasets/zongaobian/h1b-lca-disclosure-data-2020-2024/data). It contains LCA disclosure datasets from U.S. Department of Labor, covering the period from 2020 to 2024. It includes information such as case status, job title, Standard Occupational
Classification (SOC) title, location, and wages.

@@

This analysis will primarily focus on certified cases. The key research questions to be examined include:
Distribution Analysis:
What is the distribution of certified cases across various features such as SOC titles, locations, wages, and employer names?
Which SOC titles, states, and employers have the highest and lowest number of certified cases?
Variable Relationships:
How do wages vary across different states, SOC titles, and employers?
Employment Trends and Remote Work Patterns:
How have employment patterns evolved over time?
What trends can be observed in remote work prevalence before and after the pandemic?
Predictive Modeling:
Can a predictive model be developed to estimate wages based on job-related and employer-specific features?
Which factors contribute most significantly to wage determination in certified cases?

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

@@

The data will be reduced to include obeservations with certified case status and visa class of H-1B. Variable will only include RECEIVED_DATE, SOC_TITLE, EMPLOYER_NAME, EMPLOYER_STATE, WORKSITE_STATE, WAGE_RATE_OF_PAY_FROM, PREVAILING_WAGE.

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition

The dataset used in this analysis was obtained from Kaggle. The data was downloaded in CSV format.

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

The raw data import and cleaning will be conducted seprately since the dataset is large.

## Statistical analysis
_Explain anything related to your statistical analyses._
Various plots will be made to visualize the data. Linear regression will be used to analyze the relationship between variables. Machine learning methods will be used as modeling technique.

@@

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

@@

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

@@

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

@@

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
# load a few R packages
library(here)
library(knitr)
```

@@

@@

# Summary/Abstract

@@

The data is H1B LCA Disclosure Data (2020-2024) from [Kaggle](https://www.kaggle.com/datasets/zongaobian/h1b-lca-disclosure-data-2020-2024/data). It contains LCA disclosure datasets from U.S. Department of Labor, covering the period from 2020 to 2024, and includes information such as case status, job title, Standard Occupational Classification (SOC) title, location, and wages.

@@

This analysis will primarily focus on certified cases. The key research questions to be examined include:

(1)Distribution Analysis:

What is the distribution of certified cases across various features such as SOC titles, locations, wages, and employer names?

(2)Variable Relationships:

How do wages vary across different states, SOC titles, and employers?

(3)Employment Trends and Remote Work Patterns:

How have employment patterns evolved over time?

What trends can be observed in remote work prevalence before and after the pandemic?

(4)Predictive Modeling:

Can a predictive model be developed to estimate wages based on job-related and employer-specific features?

Which factors contribute most significantly to wage determination in certified cases?

@@

## Data aquisition and cleaning

The dataset used in this analysis is the H1B LCA Disclosure Data (2020-2024) from [Kaggle](https://www.kaggle.com/datasets/zongaobian/h1b-lca-disclosure-data-2020-2024/data). It comprises labor condition application (LCA) disclosure records from the U.S. Department of Labor, covering the period from 2020 to 2024. The dataset includes key information such as case status, job title, Standard Occupational Classification (SOC) title, employer details, location, and wages.

To ensure data quality, the dataset was filtered to include only certified cases and those specifically related to H-1B visa applications.

Additionally, the dataset was refined to include only the following essential variables:

RECEIVED_DATE (date the application was received); SOC_TITLE (occupational title classified under the SOC system); EMPLOYER_NAME (name of the employer); EMPLOYER_STATE (employer’s location); WORKSITE_STATE (location of the worksite); WAGE_RATE_OF_PAY_FROM (wage offered to the worker); WAGE_UNIT_OF_PAY (unit of pay values); PREVAILING_WAGE (prevailing wage for the job); PW_UNIT_OF_PAY (unit of pay values)

Observations with missing values in the selected variables were removed. All character columns except state names were standardized to lowercase.

## Statistical analysis
Various plots will be made to visualize the data. Linear regression will be used to analyze the relationship between variables. Machine learning methods will be used as modeling technique.

@@

### Univariate Analysis: Distribution Analysis

The @fig-top10_soc_title displays the 10 job titles with the highest number of applications filed between 2020 and 2024, with "Software Developers, Applications" leading at over 500,000 cases. The @fig-top10_emp_name highlights the top 10 employers by application count, with "COGNIZANT TECHNOLOGY SOLUTIONS US CORP" topping the list at approximately 93,000 cases. The @fig-top10_emp_state presents the 10 states with the most applications, with California recording over 600,000 cases. Meanwhile, the @fig-distribution_PREV_WAGE plot illustrates the distribution of prevailing wages, revealing that most cases offer wages near $100,000 per year, while fewer cases fall at the lower end (around $15,000) or approach higher values near $800,000 (see Supplement Table 1).
```{r}
#| label: fig-top10_soc_title
#| fig-cap: "Top 10 Job Titles."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_soc_title.png"))
```

```{r}
#| label: fig-top10_emp_name
#| fig-cap: "Top 10 Employers."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_emp_name.png"))
```

```{r}
#| label: fig-top10_emp_state
#| fig-cap: "Top 10 Employer Locations."
#| echo: FALSE
knitr::include_graphics(here("results","figures","top10_emp_state.png"))
```

```{r}
#| label: fig-distribution_PREV_WAGE
#| fig-cap: "Distribution of Prevailing Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","distribution_PREV_WAGE.png"))
```

### Bivariate Analysis

The @fig-wage_by_soc_title plot shows the 10 job titles with the highest median wage, and they are all medical practitioners. The @fig-wage_by_employer plot highlights the 10 employers with the highest median wage, spanning healthcare, financial services, life sciences, technology, and manufacturing. The @fig-wage_by_state plot indicates that Washington has the highest median wage by state. Plots for the 10 lowest median wages are provided in Supplement Figures 1–3. The @fig-remote_trend_absolute plot illustrates the trend of remote work over the years. The number of remote work cases surged from 2019 to 2020 due to the pandemic and remained relatively stable from 2020 to 2023. Post-pandemic, there was a slight decline from 2023 to 2024. However, as remote work patterns became more established, the decrease was not significant. This analysis uses absolute values instead of proportions because remote work is identified based on discrepancies between employer and worksite states. Some remote work cases may not be captured under this definition, potentially underestimating the actual proportion. Therefore, absolute values provide a more accurate reflection of the remote work trend.

```{r}
#| label: fig-wage_by_soc_title
#| fig-cap: "Top 10 Job Titles by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_soc_title.png"))
```

```{r}
#| label: fig-wage_by_employer
#| fig-cap: "Top 10 Employers by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_employer.png"))
```

```{r}
#| label: fig-wage_by_state
#| fig-cap: "Top 10 States by Highest Median Wage."
#| echo: FALSE
knitr::include_graphics(here("results","figures","wage_by_state.png"))
```

```{r}
#| label: fig-remote_trend_absolute
#| fig-cap: "Remote Work Trends Over Years."
#| echo: FALSE
knitr::include_graphics(here("results","figures","remote_trend_absolute.png"))
```

@@

@@

@@

## Summary and Interpretation

## Strengths and Limitations

## Conclusions
